---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Emerson College Polling Methodology (October 14-16, 2024)

## Overview
Emerson College Polling, known for its extensive online polling methods, conducted a survey from October 14 to 16, 2024, targeting 1,000 likely voters. The poll measured voter preferences between Kamala Harris and Donald Trump, revealing a near tie with Harris at 50% and Trump at 49%.

## Population, Frame, and Sample
The target population is likely voters in the U.S. elections. These voters are determined by the voters' likeliness to vote in the upcoming elections and the voter's voting history. Both of these factors are self-reported. 

The sampling frame is the voters that are on the Aristotle list and on the online panel provided by CINT. Aristotle is an online supplier of voter files and CINT is a software company that connects researchers, such as Emerson, to online panels. This approach helps ensure a representative sample by accurately reflecting the electorate.

The sample size is the 1,000 likely voters chosen arbitrarily from the sampling frame. Likely voters are determined by a combination of voter history, registration status, and demographic data. These factors are self-reported. 

## Sample Recruitment and Approach
Emerson College used a mixed-mode sampling approach for its polling. Specifically, three primary modes were used to collect data for its October 14-16, 2024 poll:

- MMS-to-web text surveys: Respondents were contacted via text messages sent to cell phones. These messages directed participants to complete the survey online. This method was conducted using voter lists provided by Aristotle.

- Interactive Voice Response (IVR): Landlines were targeted using automated phone calls where respondents could answer questions using their keypads. The contact information for this was also sourced from Aristotle’s voter lists.

- Online Panel from CINT: Emerson utilized a pre-screened, opt-in online panel of voters provided by CINT.

## Advantages and Trade-offs in Mixed-Mode Sampling
The multi-mode sampling approach used in this poll has several advantages:

- Wider demographic reach: Having a combination of data collection methods allows Emerson to gather responses from a broader demographic group. In this polling, MMS-to-web surveys mostly gather responses from younger voters and those who primarily use mobile devices. IVR target older and rural voters, who are more likely to respond via landlines. Online panels is more efficient in reaching tech-savvy individuals. 

- Cost-efficient: Automated technologies reduce the cost compared to live interviews. In a polling context, this allows for a larger sample size.

However, the trade-offs include:

- Coverage bias: IVR only reaches individuals with landlines, which may skew results towards older demographics.

- Self-selection bias: Respondents in online panels tend to be more engaged, which can lead to over-representation of politically active individuals.

## Non-response Handling
Emerson addresses non-response using a weighting system. The data is adjusted based on key demographic variables like age, gender, race, education, and party affiliation, making the sample more representative of the electorate. Weights are applied to compensate for under- or over-represented groups, ensuring balanced results.

## Questionnaire Design
The Emerson survey features a straightforward, efficient design focused on key issues and voter preferences. Strengths of the questionnaire include:

- Simplicity: Clear, concise questions makes the survey easy for respondents to understand and to provide accurate answers.

- Topical relevance: The poll zeroes in on vote preference and demographic splits, delivering timely insights during election cycles.

However, weaknesses include:

- Lack of nuance: While the survey captures basic preferences, it does not delve deeply into voters' motivations or the factors driving their choices.

- Mode limitations: Different survey modes can affect how respondents engage with the questions. This can potentially influence the depth and accuracy of voter responses.

## Conclusion
Emerson College Polling's mixed-mode methodology is effective in balancing cost with broad demographic reach. Its weighting techniques contribute to reliable and accurate results despite non-responses. While the use of multiple survey modes increases population representation, possible biases—such as lack of nuance and mode effects need to be taken into consideration when interpreting results. 

In sum, this method is a strong predictor of election outcomes and continues to adapt to the evolving challenges of modern polling.

# Idealized Methodology

## Overview
This methodology outlines an election forecasting plan with a budget of $100,000. Heavily inspired by TIPP Insights and Emerson College's hybrid approaches, the survey methodology balances cost-efficiency and data accuracy by using different modes of data collection and wide demographic reach. 

This forecasting plan will predict the popular vote and Electoral College outcomes, emphasizing key swing states and underrepresented voter groups.

## Sampling Approach
This idealized methodology uses a multi-mode hybrid sampling approach. This allows the use of both probability-based and non-probability-based methods:

- Probability-based methods:

  - Live phone interviews (landline and cell): As used by TIPP, live phone interviews is a direct and personal method of collecting responses. Voter lists would be sourced from Aristotle to ensure we contact likely voters. This mode reaches older voters and voters with less internet connection. 
  
  - Interactive Voice Response (IVR): Similar to Emerson’s approach, IVR for landline users balances costs and ensures demographic representation, particularly among older or less tech-savvy populations. This method captures quick responses and is less expensive than live interviews.

- Non-probability methods:

  - MMS-to-web (mobile): Following Emerson’s MMS-to-web strategy, we would send multimedia messages to likely voters’ cell phones, encouraging them to complete an online survey. This method works well for younger, mobile-first voters, helping fill in gaps where traditional methods might miss.
  
  - Online panel recruitment via CINT: To mirror TIPP’s online panel approach, a portion of the survey would be conducted via an opt-in panel of pre-screened voters from CINT, capturing tech-savvy respondents. CINT’s platform ensures that different voter groups, including hard-to-reach groups, are represented in the polling.

## Recruitment Strategy
By using Aristotle’s voter lists and CINT's platform as a sampling frame, the sample consists of likely voters based on registration status and past voting behavior. Quota sampling would be used to match the demographic profile of the U.S. electorate. This ensures that respondents in the sample reflect key demographic variables such as age, gender, race, and education level.

We would oversample key swing states like Pennsylvania, Wisconsin, and Florida. Key swing states are states where there is no overwhelming support for a particular candidate or political party. The outcome in these states often determine the overall result of the election, making them critical in securing a win in the Electoral College. Oversampling these states allows a more accurate forecasting in these battlegrounds, which are critical for predicting the Electoral College outcome.

## Data Validation

Given that the data is collected in a hybrid format, we would use duplicate detection methods via IP addresses or phone numbers to eliminate multiple responses from the same individual. Also, we would implement attention checks to identify respondents who are not fully engaged. Both duplicate respondents and unengaged respondents bias results, and should be removed from the data set. 

Following TIPP’s iterative weighting process, the data will be weighted to ensure that the sample aligns with key demographics (race, gender, education) based on census data.

## Conclusion
By combining TIPP’s live phone/IVR hybrid and Emerson College’s mixed-mode survey methods, this idealized methodology captures diverse voter demographics. It uses both probability-based and non-probability-based sampling methods to forecast the 2024 U.S. Presidential Election. 

A link to the survey can be found at: https://forms.gle/haGLaQBPQhrXE6Xa7 

# References
